The Free-play Sandbox: a Novel Methodology for the Evaluation of Social Robotics

### 2017-10-07 paper.tex
Reworked discussion and conclusion

\section{Discussion}
\label{sec:discussion}

\subsection{Analysis of the free play sandbox}

Our free play sandbox elicits a loosely structured play situation. The
situation is indeed essentially aimless (free play). The goals and play episodes
that we observe during the interactions are not pre-defined, and essentially
unknown until they are decided by the players. In this sense, the free play
sandbox permits and elicits what is fundamentally \emph{unstructured play}.
\todo{Is it really aim-less? Children create their own goals, no?}

However, a second look reveals some important structuring elements.

First, the physical bounds of the sandbox (an interactive table) limit the
play zone to a well defined and relatively small area. As a consequence,
children are mostly static (they are sitting in front of the table) and their
primary form of interaction is 2D pick and place of items by drag and dropping
them on the screen.

Second, the game items themselves structure the game scenarios. Items are either
iconic characters (animals or children) or plain blocks
(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
(like 'crocodiles like water and eat children'), while the later have little
semantics associated. The game background, with its recognizable zones, also
elicit a particular type of games (like building a zoo or pretending we explore
the savannah).

Two more elements (the role and place of the experimenter, and the initial
prompt given to the participant) have an important impact on the shape of the
play situation. We discuss separately these two points in the description of the
experimental procedure.

Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
actual play situations are not known and might change several times during the
interaction; the game actions, even though all based on a single modality (picking and
placing game items), are unlimited; the social interactions between participants
are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
unconstrainted. However, the broad domain of the play situations and the range of
possible social interactions is bound by the physical bounds of the play zone
and the theme of the game items.
\todo{more than one actions, also drawing}

\paragraph{Forms of play elicited by the sandbox}

\begin{itemize}
    \item dramatic play
    \item storytelling
    \item creative play
    \item pretend play
\end{itemize}

\subsection{An exploration of the dynamics of social interactions}

\begin{figure}
\centering

\resizebox{0.8\linewidth}{!}{%
\begin{tikzpicture}[
                font=\sffamily,
                >=latex,
                every edge/.style={draw, very thick},
                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]

    %%% NODES
    \node [node] (mentalrepresentations) {Mental Representations};
    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};

    %%% CONNECTIONS
    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);

\end{tikzpicture}
}
\label{fig|researchquestions}
\caption{Game dynamics, social dynamics and mental representations are linked
    with each other through social actions (implicit and explicit communication)
    and actions on the environment (mainly game actions)}
\end{figure}

The mechanisms underlying the social interactions happening during free play
raise important questions for social robotics.

\paragraph{Game Dynamics}

\begin{itemize}
    \item do we observe a sequence of sub-games?
    \item how are these sub-games initiated?
    \item how does A keep track of what B is playing at?
    \item how do we segment the action flow into meaningful play situations?
\end{itemize}

\paragraph{Social Dynamics}

\begin{itemize}
    \item does a ``social protocol'' establish?
    \item if so, what are these social rules?
    \item implicit vs explicit rule setting?
    \item what interaction modalities (or combination thereof) are relied upon
        to establish and maintain this ``social protocol''?
\end{itemize}

\paragraph{Mental representations}

\begin{itemize}
    \item Does A know what B is thinking of at time t?
    \item Do the participants perform explicit grounding of their respective mental models
        (ie, ask the other one what she is doing/thinking/planning to do?)
    \item How does the implicit grounding take place? attention tracking?
    \item Do we need at all to know what the other intends to do to actually
        collaborate? surface alignment vs deep grounding
    \item Procedural (how do you do it?) vs Semantic
        grounding (what are you doing? playing at?)
\end{itemize}

\subsection{Stages of Play}

\section{Conclusion}
\label{sec:conclusion}

We presented a set-up and data set of relatively unconstrained interaction
between children and between a robot and a child. The set-up captures a rich set
of multimodal streams which can be used to mine the social, verbal and
non-verbal communication between two parties engaging in a relatively rich free
play interaction. The data holds considerable promise for training social signal
interpretation software, such as engagement interpretation or eye gaze reading.
The dataset collected has sufficiently rich data and a wide range of multi-modal
dimensions making it particularly suitable for Deep Learning of social signal
processing algorithms. It also allow for very rich input to action selection
mechanisms needed for autonomous robot behaviour. Future work will focus on
using mining the data for patterns in free play, as per Parten's classification,
and will attempt to extract social signals relevant to drive the interaction.
Early results show that Deep Learning shows considerable promise for
high-resolution tracking of eye gaze from the RGB video streams.

>>>>

\section{Discussion \& Conclusion}
\label{sec:discussion}

\subsection{Analysis of the free-play sandbox}

The free play sandbox elicits a \emph{loosely structured} form of play: the
actual play situations are not known and might change several times during the
interaction; the game actions, even though based on a single interaction
modality (the touchscreen), are varied and unlimited (especially when
considering the drawings); the social interactions between participants
are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
unconstrained. This loose structure creates a fecund environment for children to
express a range of complex, dynamics, natural social behaviours that are not
tied to an overly constructed social situation.

The interaction is loosely structure. It is nonetheless structured:
First, the physical bounds of the sandbox (an interactive table) limit the
play area to a well defined and relatively small area. As a consequence,
children are mostly static (they are sitting in front of the table) and their
primary form of physical interaction is based on 2D manipulations on a screen.

Second, the game items themselves (visible in Figure~\ref{fig|sandbox})
structure the game scenarios. They are iconic characters (animals or children)
with strong semantics associated to them (like 'crocodiles like water and eat
children'). The game background, with its recognizable zones, also elicit a
particular type of games (like building a zoo or pretending we explore the
savannah).

These elements of structure (along with other, less important, ones) make it
possible for the free-play sandbox paradigm to retain some key properties that
makes it a practical and effective scientific tool: because the game builds on
simple and universal play mechanics (drawings, pretend play with characters),
the paradigm is essentially cross-cultural; because the sandbox is physically
bounded and relatively small, it can be easily transported and practically
deployed in a range of environments (schools, exhibitions, etc.); because the
whole apparatus is well defined and relatively easy to duplicate (it essentially
consists in one single touchscreen computer), the free-play sandbox facilitates
replication of findings in HRI while preserving ecological validity.

%\paragraph{Forms of play elicited by the sandbox}
%
%\begin{itemize}
%    \item dramatic play
%    \item storytelling
%    \item creative play
%    \item pretend play
%\end{itemize}
%
%\subsection{An exploration of the dynamics of social interactions}
%
%\begin{figure}
%\centering
%
%\resizebox{0.8\linewidth}{!}{%
%\begin{tikzpicture}[
%                font=\sffamily,
%                >=latex,
%                every edge/.style={draw, very thick},
%                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
%                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
%
%    %%% NODES
%    \node [node] (mentalrepresentations) {Mental Representations};
%    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
%    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
%    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
%    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
%
%    %%% CONNECTIONS
%    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
%    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
%    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
%    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
%    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
%    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
%    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
%
%\end{tikzpicture}
%}
%\label{fig|researchquestions}
%\caption{Game dynamics, social dynamics and mental representations are linked
%    with each other through social actions (implicit and explicit communication)
%    and actions on the environment (mainly game actions)}
%\end{figure}
%
%The mechanisms underlying the social interactions happening during free play
%raise important questions for social robotics.
%
%
%\paragraph{Game Dynamics}
%
%\begin{itemize}
%    \item do we observe a sequence of sub-games?
%    \item how are these sub-games initiated?
%    \item how does A keep track of what B is playing at?
%    \item how do we segment the action flow into meaningful play situations?
%\end{itemize}
%
%\paragraph{Social Dynamics}
%
%\begin{itemize}
%    \item does a ``social protocol'' establish?
%    \item if so, what are these social rules?
%    \item implicit vs explicit rule setting?
%    \item what interaction modalities (or combination thereof) are relied upon
%        to establish and maintain this ``social protocol''?
%\end{itemize}
%
%\paragraph{Mental representations}
%
%\begin{itemize}
%    \item Does A know what B is thinking of at time t?
%    \item Do the participants perform explicit grounding of their respective mental models
%        (ie, ask the other one what she is doing/thinking/planning to do?)
%    \item How does the implicit grounding take place? attention tracking?
%    \item Do we need at all to know what the other intends to do to actually
%        collaborate? surface alignment vs deep grounding
%    \item Procedural (how do you do it?) vs Semantic
%        grounding (what are you doing? playing at?)
%\end{itemize}
%
%\subsection{Stages of Play}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Towards the machine learning of social interactions?}
\label{sec:conclusion}

We presented a set-up and data set of relatively unconstrained interaction
between children and between a robot and a child. The set-up captures a rich set
of multimodal streams which can be used to mine the social, verbal and
non-verbal communication between two parties engaging in a rich free-play
interaction. The data holds considerable promise for training social signal
interpretation software, such as engagement interpretation or eye gaze reading.
The dataset collected has sufficiently rich data and a wide range of multi-modal
dimensions making it particularly suitable for Deep Learning of social signal
processing algorithms. It also allow for very rich input to action selection
mechanisms needed for autonomous robot behaviour. Future work will focus on
mining the data for social patterns occurring in play situations, as per
Parten's classification, and will attempt to extract social signals relevant to
drive the interaction. Some early results show, for instance,  that deep
learning shows considerable promise for high-resolution tracking of eye gaze
from the RGB video streams.
